CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility

### 2021-02-07 codecheck-paper.tex
finished up to "core principles" section

\usepackage{fontawesome}
\newcommand{\github}[1]{\href{https://github.com/#1}{\faGithub}}
\newcommand{\twitter}[1]{\href{https://twitter.com/#1}{\faTwitter}}

@@

\begin{document}
%\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
\title{CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, GB; \email{sje30@cam.ac.uk}}
\affil[$\ast$]{Contributed equally}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
\maketitle
\begin{abstract}
  In its traditional format, the scientific paper falls short of effectively communicating computational research.
  Since this issue is difficult to address due to lacking incentives, guidance, benefits, recognitions, and reproducibility, we come at this problem from a different angle:
  We propose to introduce, as part of peer review, a system by which the workflows underlying research articles are executed.
  The CODECHECK system, which uses free intrastructure and open tools, relies on five principles and can be integrated into review and publication processes in multiple ways.
  We describe these integrations along multiple dimensions (importance, who, openness, when).
  In collaboration with academic publishers and conferences, we demonstrate CODECHECK with 25 reproductions of diverse scientific publications.
  These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability.
  While CODECHECK has clear limitations, it may represent a building block in the Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
\end{abstract}

@@

Many areas of scientific research use computations to either simulate or analyse their data. These computations are becoming increasingly complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
To complement the traditional route of sharing research by writing papers,
there is a growing demand to share the underlying artefacts, notably 
code and datasets, so that others can inspect, reproduce or expand that work
(see Figure~\ref{fig:inverse}).
Some of the earliest proponents of this initiative were Buckheit and Donoho \cite{buckheit_wavelab_1995}, who coined what has been called \emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{``An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures.''}

So, assuming that researchers begin to share more artefacts, how might these artefacts be examined to ensure that they do what they claim?
For example,
most scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share.
Yet, each journal can implement this requirement differently.
At one end of the spectrum, journals have been created to accept ``data papers''
(e.g.,~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}});
these journals have established rigorous procedures by which
the data is validated according to standards in each field.
At the other end of the spectrum, many journals still simply allow authors to state ``Data available upon reasonable request''.
Authors, while possibly well intentioned at the time of writing the article, often cannot provide the data when readers ask for it, as data disappears over time \cite{Vines2014-hf}.

Given that no clear standards yet exist for sharing data, what hope might there be for sharing computer programs?
Both data and software together are required to validate outcomes of a computational analysis, but they differ in that data can be seen as static and inert,
while code requires an environment to be run in and interacted with.
This makes software harder to share.
Our anectodal experiences on this matter suggest that researchers often give a variety of reasons for why code cannot be shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
Our view, rooted in Open~Science principles\footnote{\url{https://en.wikipedia.org/wiki/Open_science}}, is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
Having code and data openly available and archived provides a valuable resource for others to learn from, even if the code no longer runs or if there is no documentation.
However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting as early as possible and that such a focused approach can greatly reduce the barrier to evaluating non-text research materials.
Just as data journals' validations of data and all journals' peer review can provide a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, we think the same baseline should be provided for the workflow underlying a paper.
With this in mind, we have developed a set of principles and an example workflow that provides what we hope is a pragmatic way of checking that a paper's code works.

As this work's main contribution, we offer a thorough description of a process and its variations to integrate a much needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across different scientific disciplines.
We call this system CODECHECK as outlined next.

@@

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the AUTHOR of a paper providing the code to be checked,
(2) the PUBLISHER of a journal interested in publishing the author's paper, and
(3) the CODECHECKER, who checks that the authors's code works.
The workflow we have refined is shown in Figure~\ref{fig:workflow}
and consists of six steps.

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
% emoji codes:
% book stack: 1f4da
% package: 1f4e6
% file drawer: 1f5c4
% disk: 1f4be
% document: 1f4c4
% phd: 1f393
% detective: u1f575
% laptop: 1f4bb
% folder: 1f4c1 1f4c2
% receipt: 1f9fe
  \centering
      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
  \caption{The CODECHECK example process implementation. Codechecker act as a detectives:
  They investigate and record, but do not fix issues.}
  \label{fig:workflow}
\end{figure}

@@

\textbf{Step 2:} The publisher finds a CODECHECKER to check the code. This is
analogous to the publisher finding one or more reviewers to evaluate
the paper.
However, a crucial difference is that we suggest the CODECHECKER and the author are able to talk freely and directly to each, regardless how open the scientific peer review process is.

@@

\textbf{Step 4:} The CODECHECKER writes a certificate stating how the code was
run and includes a copy of outputs (e.g.,~data, figures, or tables) that were
independently generated.
The certificate may include further recommendations to the author on how to improve the material.
The free text in the certificate is the most flexible and feasible way to 
make clear what was really checked, because each workflow is unique.
Since no specific tool or platform should be required, such that no authors are excluded, it would be futile for the CODECHECKER to use automation or fixed checklists.

\textbf{Step 5:} The certificate as well as auxiliary files created
during the check, e.g.,~a specification of a computing environment, data 
subsets or helper scripts, and the original code and data get deposited in
an open archive unless restrictions regarding data size, license or 
sensitivity apply.
Currently, the CODECHECKERs deposit the material on Zenodo themselves, but the PUBLISHER may complete this step after integrating CODECHECK into its review process.
A badge or other visual aid may be added to the deposit and the paper and
link to the certificate.
Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledgement of the completed CODECHECK.
The badge and the actual check are incentives for undertaking the effort needed to provide a readily computationally reproducible workflow.

\textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
certificate to scientific reviewers or handling editors or publish it
and link between the certificate, the paper, and any repositories 
through metadata. Currently, the CODECHECKER creates these connections on
Zenodo\footnote{
See ``related identifiers'' and ``alternate identifiers'' in the right-hand
sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
The PUBLISHER also gives credit to the CODECHECKER for their work by depositing
the activity appropriately in scholarly profiles, such as ORCID\footnote{
See peer review contributions in ORCID records at
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
certificate repository to the published paper or the original code repository.

@@

The example workflow leaves room for this process to be implemented differently.
To elaborate on these options, we consider several dimensions in a space of possible CODECHECK workflows, as shown in Figure~\ref{fig:dimensions}, to represent relevant aspects for the involved stakeholders.
These aspects touch on timing, responsibilities, and 
transparency and will be discussed in the following sections.
The options on each dimension have their own pros and cons and can,
for the most part, be freely combined.

@@

The point in time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
The earlier a CODECHECK happens in the publishing process, the more it can affect editorial decisions:
Is a paper published, sent back for revisions, or rejected?
Even earlier checks, i.e.,~a CODECHECK of a preprint, may even help to improve the workflow itself, even before a publisher is involved.
As such, codechecking papers could be part of a preprint server's policy or simply be initiated by interested readers and community members.

Publishers could introduce a CODECHECK as a \textbf{strict prerequisite}.
As this can reduce the workload of reviewers, such a check should not occur late in the review process.
Yet, the later in the review process the check happens, the easier is it to allow bidirectional communication between the author and codechecker, e.g., because the author might already be notified of the paper's acceptance and may more freely share materials online closer to the paper's publication date.
A \textbf{pre-review} CODECHECK would allow editors to only send a submission out for peer review if it passes the check, or to include the certificate in the submission package provided to the reviewers.
Reviewers may then judge the the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.

A CODECHECK may also be conducted in \textbf{parallel} to the scientific peer review.
This puts less burden on the turnaround time for the check, yet it only makes the outcomes available during the final consideration by the handling editor.
The check could also be assigned after suggestion by a reviewer, which would remove the need for submissions to undergo a pre-review screening.
However, soliciting such a ``specialist review'' is much less desirable than having a regular CODECHECK, thus avoiding the situation in which some submissions get special treatment.
In both cases, a CODECHECK could then be awarded \textbf{equal weight} compared to the other reviews for the editor's decision.

A \textbf{post-acceptance} CODECHECK would have the smallest impact on editorial decisions and may simply provide \textbf{extra merit} on top of the submission's acceptance.
This is the least impactful solution in which all material is still evaluated and the results of the check are properly acknowledged, because the check can be completed before the publication of the paper.
The GIScience group of checks (see below) falls into this category:
By displaying a badge on the volume and article landing pages, the AGILE conference highlights articles whose reproducibility was successfully reviewed.
Similarly, in collaborations with journals, some GIScience have been checked while the authors worked on revisions.

A CODECHECK may also be conducted \textbf{post-publication}, though it would require someone to update the article and article metadata to reference the check; otherwise, readers would presumably not be aware of it.
In general, publishers hesite to make such revisions to published articles.
While this option has the least impact on current publishing practices and enables open discourse, it also has the smallest effect on the original submission and does not acknowledge the importance of reproducible workflows for ensuring good scientific practice.

In general, the timing of a CODECHECK needs to be connected with 
established or typical scholarly communication process.
While more radical ideas do exist for changing the paradigms of how researchers can more granually share their work much
\footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
% mimosa is another platform idea, but no real prototype AFAICT:
% https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
% https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
Enhancing existing processes with CODECHECKs allows communities to transition towards more open practices at their own pace and ensure that no one is unfaily left behind.
When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
Depending on when and who conducts the check, it might be done quickly or it might add considerable time to how long it takes from submission to publication.
In our first checks, we found that a CODECHECK generally takes from 2-5 hours, with some outliers on the higher end.
This time includes writing and publishing the certificate but excludes the time for actual computations.
These efforts are comparable to the time needed to peer review a submission,
which aligns with the efforts some volunteer codecheckers are willing to
make.
However, the time includes a considerable amount of communicating about the
process, especially regarding who publishes which document when, so that proper
cross-referencing between paper and certificate is ensured via persistent
identifiers.
When integrated into a peer review platform, this handling of documents should
become much more effective and reduce organisational overhead.

@@

The topic of anonymity is broadly discussed, especially in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stance on this topic, the general motivation 
behind CODECHECK
for higher transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can also serve to protect individuals \cite{tennant_limitations_2020}, e.g.,~junior scientists, whereas revealing of names from tenured researchers might be more reasonable to require.
The possible negative effects of a non-anonymous review are reduced if a CODECHECK is not relevant for a journal's decision to accept or reject, but that is, of course, not desirable when the goal is higher transparency and reproducibility.
Instead, we argue that a CODECHECK is technical process that should in general find fixable problems; it is not aimed at giving an opinion or identifying a faulty approach.
It is technically possible for a nonsense workflow to receive
a CODECHECK certificate.
If passing a CODECHECK becomes mandatory, the full transparency may have to be 
revisited as the relations between authors and codecheckers would fall under the
same social and community challenges as non-anonymous peer review
(cf.~\cite{everythinghertz123}).
% previous sentenc inspired by https://osf.io/9cftx/

The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in allowing for bidirectional communication between author and codechecker.
Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the documentation or the update code.
Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
Of course, the codechecker may also provide a fix within the code instead
of lengthily describing a small change if it can be quickly done.
Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
While communication between author and codechecker may be facilitated in an 
anonymous way by the 
publisher, it most likely only helps to protect the identity of the codechecker, because, especially when codebases are of a certain size, not trivial, or if very specialised, they are almost impossible to fully anonymise.
Therefore, the most effective and desirable situation for the stakeholders is to hold a non-anonymous and collaborative CODECHECK.
The contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit can be given within the research software development community.
Nevertheless, depending on the needs represented by other dimensions,
the aspect of ``Who knows who?'' may be the most flexible.

@@

The options for who can conduct a CODECHECK require a somewhat complex matching of a combination of skills and availability.
Ideally, thecodechecker has a matching code \emph{and} domain expertise; for example, to check a workflow based on Python code analysing neuroimaging datasets, the codechecker should ideally be experienced in Python and neuroimaging analysis.
However,
a well-documented workflow should be executable by any person with a basic
understanding of computers. Naturally, the more prerequisite knowledge the
codechecker has, the less time is needed to understand the goals and 
the mechanics of an analysis.
From our experiences, the priority should be given to matching technical expertise first, as lacking knowledge in setting up a computing environment with a particular language or tool is much more of a problem than assessing the outcome, e.g.,~comparing created figures with the original, without an in-depth understanding of the science.
The depth of the check will mostly be driven by the time required and expertise of the checker, though in general, we expect a CODECHECK not to evaluate performance (e.g.,~it will not check whether a racing car has the best apex speed) but to assess operation free of failure (i.e.,~is it a car with 4 wheels and does it move forwards when accelerating).
This depth can be seen independent of the check's importance, as long as it is communicated clearly to all stakeholders, most importantly authors and readers.

The concrete people to provide such expertise can range from researchers,
possibly drawn from the regular pool of volunteer
\textbf{scientific reviewers}, or from a special group of
\textbf{reproducibility reviewers} via specific roles
such as \textbf{reproducibility editors}, to experts working as part of
\textbf{editorial staff} with a publisher. If the regular reviewers
are relied upon, the editor's task to find suitable ones becomes more
tedious; extra information such as software familiarity would have to be
modelled in review platforms.

We consider a \emph{single codechecker} to be enough, unlike peer review where
multiple referees are common.
Because of the technical focus, determining whether the concrete outputs produced by the author and the codechecker, who can collaborate to solve any issues, actually match should be a very factual process and not require much debate or opinion.
Code usually harbours systematic and repeatable mistakes and is thereby more reliable and auditable than processes controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.
Depending on the importance of the CODECHECK
(see above), a mediator, e.g.,~the editor, or a
second opinion might be warranted in rare cases.

>>>>

\usepackage{fontspec}
\usepackage{fontawesome}
\newcommand{\github}[1]{\href{https://github.com/#1}{\faGithub}}
\newcommand{\twitter}[1]{\href{https://twitter.com/#1}{\faTwitter}}

@@

\begin{document}
%\title{CODECHECK: An open-science initiative for the independent execution of computations underlying research articles}
\title{CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review improves reproducibility}
\author[1,$\ast$]{Daniel~N\"{u}st \orcid{0000-0002-0024-5046} \github{nuest} \twitter{nordholmen}}
\author[2,$\ast$]{Stephen~J~Eglen \orcid{0000-0001-8607-8025} \github{sje30} \twitter{StephenEglen}}
\affil[1]{Institute for Geoinformatics, University of M\"unster, Germany; \email{daniel.nuest@uni-muenster.de}}
\affil[2]{Department of Applied Mathematics and Theoretical Physics, University of Cambridge, UK; \email{sje30@cam.ac.uk}}
\affil[$\ast$]{Contributed equally}
\affil[$$]{Public project for this article is at \url{https://github.com/codecheckers/paper}.}
\maketitle
\begin{abstract}
  The traditional scientific paper falls short of effectively communicating computational research.
  Since this issue is difficult to address due to lacking incentives, guidance, benefits, recognitions, and reproducibility, we come at this problem from a different angle:
  We propose a system by which the computational workflows underlying research articles are checked.
  The CODECHECK system uses open intrastructure and tools and can be integrated into review and publication processes in multiple ways.
  We describe these integrations along multiple dimensions (importance, who, openness, when).
  In collaboration with academic publishers and conferences, we demonstrate CODECHECK with 25 reproductions of diverse scientific publications.
  These CODECHECKs show that asking for reproducible workflows during a collaborative review can effectively improve executability.
  While CODECHECK has clear limitations, it may represent a building block in Open Science and publishing ecosystems for improving the reproducibility, appreciation, and, potentially, the quality of non-textual research artefacts.
\end{abstract}

\clearpage

@@

Many areas of scientific research use computations to simulate or analyse their data. These computations are complex and are difficult to explain coherently in a paper \citep{marwick_how_2015}.
To complement the traditional route of sharing research by writing papers,
there is a growing demand to share the underlying artefacts, notably 
code and datasets, so that others can inspect, reproduce or expand that work
(see Figure~\ref{fig:inverse}).
Some of the earliest proponents of this initiative were Buckheit and Donoho \cite{buckheit_wavelab_1995}, who coined what has been called \emph{Claerbout's claim} (extending on \citet{claerbout_electronic_1992}):
\emph{``An article about computational science in a scientific publication 
is \text{not} the scholarship itself, it is merely \textbf{advertising} of
the scholarship. The actual scholarship is the complete software development
environment and the complete set of instructions which generated the 
figures.''}

If researchers start sharing more artefacts, how might these artefacts be examined to ensure that they do what they claim?
For example,
although  scientific journals now require a data sharing statement that
outlines what data the authors have (or will) share, journals  implement this differently.
At one end of the spectrum, journals have been created to accept ``data papers''
(e.g.,~\href{https://www.nature.com/sdata/}{\emph{Scientific~Data}}, 
\href{https://essd.copernicus.org/}{\emph{Earth System Science Data}}, 
\href{https://rmets.onlinelibrary.wiley.com/journal/20496060}{\emph{Geoscience Data Journal}},
\href{https://bdj.pensoft.net/}{\emph{Biodiversity Data Journal}},
\href{https://openpsychologydata.metajnl.com/}{\emph{Journal of Open Psychology Data}},
\href{https://odjar.org/}{\emph{Open Data Journal for Agricultural Research}},
\href{https://openhealthdata.metajnl.com}{\emph{Journal of Open Health Data}});
these journals have established rigorous procedures by which
the data is validated according to standards in each field.
At the other end of the spectrum, many journals still allow authors to state ``Data available upon reasonable request''.
Authors, while possibly well intentioned at the time of writing the article, often cannot provide data when requested as data disappears over time \cite{Vines2014-hf}.

Given that no clear standards yet exist for sharing data, what hope might there be for sharing computer programs?
Both data and software are required to validate a
computational analysis; data can be seen as inert whereas
code requires an environment to be run in.
This makes software harder to share.
Our experience is that researchers offer several reasons for why code
is not shared, e.g.,~``there is no documentation'', ``I cannot maintain it'', or ``I do not want to give away my code to competitors''.
Our view is that sharing code, wherever possible, is good for the community and the individual \cite{Barnes2010-iv,markowetz_five_2015}.
Having code and data openly available and archived provides a valuable
resource for others to learn from, even if the code is broken or lacks documentation.
However, with a little effort, we believe that if an independent person can execute an analysis, this is worth documenting and that this  reduces the barrier to evaluating non-text research materials.
Just as data journals' validations of data and all journals' peer review  provides a ``baseline reassurance'', i.e.,~that a paper has been checked by someone with an understanding of the topic \cite{fyfe_mission_2019}, the same baseline could be provided for the workflow underlying a paper.
With this in mind, we have developed a set of principles and an example workflow that provides a pragmatic way of checking that a paper's code works.

Here we offer a thorough description of a process and its variations to integrate a much-needed evaluation of computational reproducibility\footnote{Following the Claerbout/Donoho/Peng terminology, see \cite{barba_terminologies_2018}.} into peer review, and we demonstrate its feasibility by means of 25 reproductions across scientific disciplines.
We call this system CODECHECK.

@@

CODECHECK is best demonstrated by way of our example workflow, and later
we expand on the underlying principles. The workflow involves three
groups of people:
(1) the AUTHOR of a paper providing the code to be checked,
(2) the PUBLISHER of a journal interested in publishing the author's paper, and
(3) the CODECHECKER, who checks that the authors's code works.
The six-step workflow we have refined is shown in Figure~\ref{fig:workflow}.

\begin{figure}
% original figure source: https://docs.google.com/drawings/d/1tGOZFNZle-oE1Ynw_tLZFmNcv3wgxD0HPs5bXzDOrng/edit
% emojis: https://github.com/googlefonts/noto-emoji and https://github.com/twitter/twemoji
% emoji codes:
% book stack: 1f4da
% package: 1f4e6
% file drawer: 1f5c4
% disk: 1f4be
% document: 1f4c4
% phd: 1f393
% detective: u1f575
% laptop: 1f4bb
% folder: 1f4c1 1f4c2
% receipt: 1f9fe
  \centering
      \includegraphics[width=\textwidth]{figs/codecheck_overview.pdf}
      \caption{The CODECHECK example process implementation.
        Codechecker act as a detectives: They investigate and record,
        but do not fix issues.  Numbers in bold refer to the six steps
        outlined in the text.}
  \label{fig:workflow}
\end{figure}

@@

\textbf{Step 2:} The publisher finds a CODECHECKER to check the
code. This is analogous to the publisher finding one or more
peer-reviewers to evaluate the paper, except we suggest that the
CODECHECKER and the author talk directly to each other.

@@

\textbf{Step 4:} The CODECHECKER writes a certificate stating how the
code was run and includes a copy of outputs (e.g.,~data, figures, or
tables) that were independently generated.  The certificate may
include further recommendations on how to improve the material.  The
free text in the certificate can describe what was really checked,
because each workflow is unique.  Since no specific tool or platform
should be required, such that no authors are excluded, it would be
futile for the CODECHECKER to use automation or fixed checklists.

\textbf{Step 5:} The certificate and auxiliary files created
during the check, e.g.,~a specification of a computing environment, data 
subsets or helper scripts, and the original code and data get deposited in
an open archive unless restrictions regarding data size, license or 
sensitivity apply.
Currently, the CODECHECKERs deposit the material on Zenodo themselves, but the PUBLISHER may complete this step after integrating CODECHECK into its review process.
A badge or other visual aid may be added to the deposit and the paper and
link to the certificate.
Although a badge simplifies the CODECHECK into a binary value and risks introducing confusion regarding the extent of the check, especially across disciplines and methods, a badge can provide recognition value and acknowledgement of the completed CODECHECK.
The badge and the actual check are incentives for undertaking the effort needed to provide a readily computationally reproducible workflow.

\textbf{Step 6:} The PUBLISHER can, depending on the timing, provide the
certificate to peer-reviewers or handling editors or publish it
and link between  certificate, paper, and any repositories 
through metadata. Currently, the CODECHECKER creates these connections on
Zenodo\footnote{
See ``related identifiers'' and ``alternate identifiers'' in the right-hand
sidebar of certificate \texttt{2020-025} \cite{cert-2020-025}.}.
The PUBLISHER also credits the CODECHECKER for their work by depositing
the activity appropriately in scholarly profiles, such as ORCID\footnote{
See peer review contributions in ORCID records at
\url{https://support.orcid.org/hc/en-us/articles/360006971333-Peer-Review}.}.
The PUBLISHER also ensures proper publication metadata, e.g.,~links from the 
certificate repository to the published paper or the original code repository.

@@

The example workflow leaves room for variations to this process.
Here we consider several dimensions in a space of possible CODECHECK workflows(Figure~\ref{fig:dimensions}), to represent relevant aspects for stakeholders.
These aspects touch on timing, responsibilities, and 
transparency and will be discussed in the following sections.
The options on each dimension have their own pros and cons and can,
for the most part, be freely combined.

@@

The time at which a CODECHECK is done and its ascribed importance are closely connected, so we describe the dimensions \emph{When} and \emph{Importance} together.
The earlier a CODECHECK happens in the publishing process, the more it can affect editorial decisions:
Is a paper published, sent back for revisions, or rejected?
Even earlier checks, i.e.,~a CODECHECK of a preprint, may help to improve the workflow itself, even before a publisher is involved.
As such, codechecking papers could be part of a preprint server's
policy or initiated by interested authors.

Publishers could introduce a CODECHECK as a \textbf{strict prerequisite}.
As this can reduce the workload of reviewers, such a check should occur early in the review process.
Yet, the later in the review process the check happens, the easier is it to allow bidirectional communication between the author and codechecker, e.g., because the author might already be notified of the paper's acceptance and may more freely share materials online closer to the paper's publication date.
A \textbf{pre-review} CODECHECK means editors could send a submission
for peer review only if it passes the check, or include the certificate in the submission package provided to peer-reviewers.
Reviewers may then judge the relevance of the computations for the results of the work, or follow journal guidelines regarding how strictly to apply the check's outcomes in their evaluation.

A CODECHECK may also be conducted in \textbf{parallel} to the scientific peer review.
This puts less burden on the turnaround time for the CODECHECK, yet it only makes the outcomes available during the final consideration by the handling editor.
The check could also be assigned after suggestion by a reviewer, which would remove the need for submissions to undergo a pre-review screening.
However, soliciting such a ``specialist review'' is much less
desirable than having a regular CODECHECK, thus avoiding the situation
in which some submissions get special treatment. In both cases, the
editor's decision could be based both on CODECHECK and peer-review reports.

A \textbf{post-acceptance} CODECHECK would have the smallest impact on
editorial decisions and may simply provide \textbf{extra merit} on top
of the submission's acceptance.  This is the least impactful solution
in which all material is still evaluated and the results of the check
are properly acknowledged, because the check can be completed before
publication of the paper.  The GIScience checks (see
below) falls into this category: By displaying a badge on the volume
and article landing pages, the AGILE conference highlights articles
whose reproducibility was confirmed.  Similarly, in
collaborations with journals, some GIScience articles were
checked whilst authors worked on revisions.

A CODECHECK may also be conducted \textbf{post-publication}, though
this requires an update to the article and article metadata to
reference the check so that readers can find the CODECHECK.  In
general, publishers hesitate to make such revisions to published
articles.  We do not prefer this option as it has the least impact on
current publishing practices and downplays the importance
of reproducible workflows for ensuring good scientific practice.

%The timing of a CODECHECK needs connecting to typical scholarly communication process.
%While ideas exist for radically changing how researchers can share their work
% \footnote{For example Octopus, \url{https://science-octopus.org/about}, or Hypergraph, \url{https://www.libscie.org/hypergraph}.},
% % mimosa is another platform idea, but no real prototype AFAICT:
% % https://projects.invisionapp.com/share/E9Z3RKE7W3P#/screens/436978924
% % https://docs.google.com/presentation/d/1V_K8hghgnvGfEtW7TdwtowTveZyy4Qs8rMemNyHx_Dg/edit#slide=id.p
% an evolutionary approach has value, and we do not yet suggest to completely eliminating research papers as a way of sharing work.
Enhancing existing processes with CODECHECKs allows communities to
gradually transition towards more open practices.
When integrating a CODECHECK into existing review and publication processes, the \emph{turnaround time} is crucial.
Depending on when and who conducts the check, it might be done quickly
or it might delay publication.
We found that a CODECHECK generally takes  2--5 hours, with some outliers on the higher end.
This time includes writing and publishing the certificate but excludes
actual computation time, some of which took days.
These efforts are comparable to the time needed to peer review a submission,
which aligns with the efforts some volunteer codecheckers are willing to
make.
Currently, there is considerable amount of communicating about the
process, especially regarding who publishes which document when, so that proper
cross-referencing between paper and certificate is ensured via persistent
identifiers.  When integrated into a peer review platform, this
handling of documents should become much more streamlined.

@@

Anonymity is broadly discussed, especially in the push towards
open peer review as part of the Open Science movement 
(cf. \cite{ross-hellauer_guidelines_2019}).
Without taking a strong stance on this topic, our motivation 
behind CODECHECK
for higher transparency and reproducibility does indeed favour a more open 
review process.
However, anonymity can protect individuals
\cite{tennant_limitations_2020}, e.g.,~junior scientists.  
The negative effects of a signed review may be reduced if a CODECHECK is not relevant for a journal's decision to accept or reject, but that is, of course, not desirable when the goal is higher transparency and reproducibility.
Instead, CODECHECK is a technical process that should generally find fixable problems; it is not aimed at giving an opinion or identifying a faulty approach.
If passing a CODECHECK becomes mandatory, full transparency may need
revisiting as the relations between authors and codecheckers would fall under the
same social and community challenges as open peer review
(cf.~\cite{everythinghertz123}).
% previous sentenc inspired by https://osf.io/9cftx/

The technical nature of the check and the challenge of providing sufficient documentation is why we see great benefits in bidirectional communication between author and codechecker.
Instead of trying to
fix problems or guess the next step, the codechecker can ask the author to 
rework the documentation or update code.
Instead of struggling to provide perfect instructions and as a result possibly not sharing any code or data, the author can make a best effort to document sufficiently.
Authors and readers can profit from a codecheckers' experience and approach, as during the check they may create useful and instructive files, e.g.,~a machine-readable computing environment specification.
While communication between author and codechecker may be anonymised via
the publisher, it most likely only helps to protect the identity of
the codechecker, because code is hard to anonymise.
Therefore, the most effective and desirable situation for the
stakeholders is to hold a open and collaborative CODECHECK.
The contributions by the codechecker may even be integrated into
the code of the workflow and be acknowledged as code commits. This way, 
proper credit can be given within the research software development community.

@@

Just as with peer-reviewers, a potential codechecker should have the
right skills and availability to do the work.  Ideally, the
codechecker has a matching code \emph{and} domain expertise to the
paper, although a well-documented workflow should be executable by any
computationally-competent person. Naturally, the more
prerequisite knowledge the codechecker has, the less time is needed to
understand the goals and the mechanics of an analysis.  From our
experiences, the priority should be given to matching technical
expertise first, as lacking knowledge in setting up a computing
environment with a particular language or tool is much more of a
problem than assessing the outcome, e.g.,~comparing created figures
with the original, without an in-depth understanding of the science.
The depth of the check will mostly be driven by the time required and
expertise of the checker, though in general, we expect a CODECHECK 
to consider reproducbility of the results above performance of the code.

Codecheckers could be drawn from a regular pool of volunteer
\textbf{scientific reviewers}, or from a special group of
\textbf{reproducibility reviewers} via specific roles such as
\textbf{reproducibility editors}, or \textbf{editorial staff} with a
publisher.  A \emph{single codechecker} is sufficient to verify the
workflow since it is mostly a factual process and not require much
debate or opinion.  Code usually harbours systematic and repeatable
mistakes and is thereby more reliable and auditable than processes
controlled by humans \cite{tibav:42484}, e.g.,~in a laboratory.  If
however publication of the paper depends on the CODECHECK, a second
opinion may be required.

We also see a great opportunity to involve \emph{early-career researchers} (ECRs) as codecheckers.
ECRs arguably have a high interest in learning
about the latest tools and technologies, as they build up their own
expertise and specialisation.
CODECHECK offes a way for ECRs to gain insights into new research and
highlight the importance of reproduction.  
\emph{ReScience X}, a journal devoted to reproduction and replication experiments \cite{roesch_new_2020}, shares an interest in this combination.
ECRs are also often familiar with new technologies, thus also making them likely to author CODECHECK-ready manuscripts
\footnote{Cf. the 77\% of 141 registered reports that were submitted to leading Neuroscience journals by ECRs \cite{chambers_registered_2019}.}.
As ECRs are introduced to peer review as codecheckers, they may transition into the role of scientific reviewer over time.
Overall, we see several opportunities and benefits to setting up a new process for CODECHECKING with a clear commitment to openness and transparency, independent of the current peer review process (see \emph{Openness} dimension).
